---
title: "Testing the Sowell Success Hypothesis: Do Family Values Trump Demographics?"
subtitle: "Analyzing Educational and Economic Outcomes Using US Census PUMS Data"
authors: 
  - name: "Dan Swart, CPA (ret)"
  - name: "Claude Sonnet 4.5"
date: today
date-format: long
# bibliography: manual-refs.bib
format:
  html:
    resources:
      - reference-backlinks.js
    include-after-body:    
      - text: |
          # <script type="text/javascript" src="reference-backlinks.js"></script>
    default: true         
    code-copy: true
    code-link: true        # This adds individual buttons
    code-fold: true        # Hide code by default, show on click
    code-summary: "Show the code"
    code-overflow: wrap
    code-block-bg: "#FAEBD7"
    code-block-border-left: "#31BAE9"
    embed-resources: false  # Fast for drafts. Override for final output
    include-in-header:
      - text: 
          <link href="https://fonts.googleapis.com/css2?family=Lato&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Fira+Mono&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Montserrat&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Cabin&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Roboto+Slab&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Schoolbell&display=swap" rel="stylesheet">
      - header.html
    css:
      - swart.css
      - tachyons.min.css
      - r-colors.css
    fontsize: 18pt
    lightbox: true
    page-layout: full
    fig-width: 12
    fig-height: 10
    fig-dpi: 300
    html-math-method: katex
    df-print: paged
    toc: true
    toc-float: true
    citeproc: true
    link-citations: true
    linestretch: 1.0
    
    
    
  typst:
    fig-width: 12
    fig-height: 10
    fig-dpi: 300
    margin:
      x: 1in
      y: 1in
    toc: true
    fontsize: 16pt
    mainfont: "Cabin"
  

    
  revealjs:
    slide-number: true
    transition: fade
    code-overflow: wrap
    center: true
    smaller: true
    scrollable: true
    chalkboard: true
    multiplex: false
    theme: solarized
    reference-location: margin
    logo: img/red-cross-640-435.png
    footer: "Footer text"
    code-block-height: 650px



  # docx:
  #   highlight-style: github
  #   fig_caption: true



editor: source

quarto:
  render:
    cache-refresh: true


# for .qmd filesd
execute:
  echo: true
  message: false
  warning: false
  eval: true
  fig-width: 12
  fig-height: 10


# for .rmd files
knitr:
  opts_chunk:
    echo: true
    error: false
    warning: false
    message: false
    cache: false


---


```{r}
#| label: setup
#| include: false


# install.packages(c("mapview", "survey", "srvyr", "arcgislayers"))

# census_api_key("95496766c51541ee6f402c1e1a8658581285b759", install = TRUE, overwrite = TRUE)


# # load libraries - NOT NEEDED


# Force dplyr's select to take precedence
select <- dplyr::select
filter <- dplyr::filter

# Options
base::options(scipen = 999)
base::options(qic.clshade = TRUE)
base::options(qic.linecol = 'black')
base::options(qic.signalcol = "firebrick")
base::options(qic.targetcol = "purple")
base::options(DT.options = base::list(dom = 'pBlfrti'))
base::options(shiny.maxRequestSize = 50 * 1024^2)
base::options(tigris_use_cache = TRUE)
base::options(device = "RStudioGD") 


# Flextable defaults:
flextable::set_flextable_defaults(
  font.size = 14, 
  font.family = "Cabin",
  font.color = "black",
  table.layout = "fixed",
  border.color = "darkgray",
  padding.top = 3, padding.bottom = 3,
  padding.left = 4, padding.right = 4,
  line_spacing = 1.3,
  digits = 2,
  decimal.mark = ".",
  big.mark = " ",
  na_str = "<na>",
  layout = "autofit"
)


# Set global theme for consistent plots
ggplot2::theme_set(
  ggplot2::theme_minimal(base_size = 20) +
    ggplot2::theme(
      plot.title.position = "plot",
      plot.title = ggtext::element_textbox_simple(
        family = "Cabin",
        face = "bold",
        color = "darkgreen",
        size = 26,
        fill = "yellow",
        lineheight = 1.2,
        padding = ggplot2::margin(5.5, 5.5, 0.0, 5.5),
        margin = ggplot2::margin(0, 0, 5.5, 0)
      ),
      plot.subtitle = ggtext::element_textbox_simple(
        family = "Cabin",
        color = "darkgreen",
        face = "bold",
        size = 24,
        fill = "yellow",
        lineheight = 1.2,
        padding = ggplot2::margin(5.5, 5.5, 5.5, 5.5),
        margin = ggplot2::margin(10.5, 0, 5.5, 0)
      ),
      plot.caption = ggtext::element_markdown(
        family = "Cabin",
        size = 22,
        hjust = 1,
        color = "darkblue",
        face = "italic",
        fill = "yellow",
        lineheight = 1.0
      ),
      axis.text.x = ggtext::element_markdown(
        family = "Cabin",
        face = "bold",
        color = "blue",
        size = 20,
        angle = 45,
        hjust = 1
      ),
      axis.title.x = ggtext::element_markdown(
        family = "Cabin",
        face = "bold",
        color = "blue",
        size = 20),
      axis.text.y = ggtext::element_markdown(
        family = "Cabin",
        face = "bold",
        color = "blue",
        size = 20,
        angle = 45,
        hjust = 1
      ),
      axis.title.y = ggtext::element_markdown(
        family = "Cabin",
        face = "bold",
        color = "blue",
        size = 20),
      strip.text = ggtext::element_markdown(
        family = "Cabin",
        color = "black",
        size = ggplot2::rel(1.1),
        face = "italic",
        margin = ggplot2::margin(2, 0, 0.5, 0, "lines")
      ),
      axis.text = ggtext::element_markdown(
        family = "Cabin",
        color = "black"),
      panel.background = ggplot2::element_rect(fill = "white", color = NA),
      plot.background = ggplot2::element_rect(fill = "white", color = NA),
      legend.position = "none",
      panel.spacing.x = grid::unit(1.5, "cm"),
      panel.spacing.y = grid::unit(1.5, "cm"),
      plot.margin = ggplot2::margin(20, 20, 20, 20, "pt")
    )
)


  
# Set seed for reproducibility
base::set.seed(123)

```


## Research Hypothesis

**Central Thesis**: High-performing individuals share certain behavioral characteristics (graduating high school, marrying before having children, maintaining employment) that transcend race, poverty, language barriers, and family structure. Success is primarily determined by family attitudes toward education and personal responsibility rather than demographic factors.

**Thomas Sowell's Framework**: Individuals who (1) graduate high school, (2) wait until marriage to have children, and (3) maintain employment have very high probability of success regardless of race, economic background, or ethnic origin.

## Data Sources and Methodology

### Primary Data Sources
- **PUMS (Public Use Microdata Sample)**: Individual/household level data from American Community Survey
- **ACS 5-Year Estimates**: Community-level demographic data via tidycensus
- **Geographic Scope**: 
  - Phase 1: Schertz-Cibolo-Universal City ISD, TX
  - Phase 2: San Antonio MSA, TX
  - Future phases: Controlled geographic expansion

```{r all-purpose-cache-function}
#| label: all-purpose-cache-function
#| eval: true

# Generic tidycensus caching function
cache_census_api_call <- function(
    type = base::c("acs", "pums", "variables", "decennial"),
    year = 2023,
    survey = "acs5",
    state = NULL,
    geography = NULL,
    table = NULL,
    variables = NULL,
    cache_dir = "data",
    force_refresh = FALSE
) {
  
  type <- base::match.arg(type)
  
  # Build descriptive cache filename
  state_str <- if (base::is.null(state)) {
    "none"
  } else if (base::length(state) == 1 && state == "all") {
    "all"
  } else {
    base::paste(base::sort(state), collapse = "-")
  }
  
  vars_str <- if (!base::is.null(variables)) {
    base::paste0("_", base::substr(rlang::hash(variables), 1, 8))
  } else if (!base::is.null(table)) {
    base::paste0("_", table)
  } else {
    ""
  }
  
  geo_str <- if (!base::is.null(geography)) {
    base::paste0("_", geography)
  } else {
    ""
  }
  
  cache_file <- base::file.path(
    cache_dir,
    base::paste0(type, "_", survey, "_", year, "_", state_str, geo_str, vars_str, ".rds")
  )
  
  # Create cache directory if needed
  if (!base::dir.exists(cache_dir)) {
    base::dir.create(cache_dir, recursive = TRUE)
  }
  
  # Return cached data if exists and not forcing refresh
  if (base::file.exists(cache_file) && !force_refresh) {
    base::message("Loading from cache: ", cache_file)
    base::return(base::readRDS(cache_file))
  }
  
  # Otherwise fetch from API
  base::message("Fetching from API...")
  
  result <- base::switch(
    type,
    "variables" = tidycensus::load_variables(year, survey),
    "acs" = tidycensus::get_acs(
      geography = geography,
      table = table,
      variables = variables,
      state = state,
      year = year,
      survey = survey
    ),
    "pums" = tidycensus::get_pums(
      variables = variables,
      state = state,
      year = year,
      survey = survey
    ),
    "decennial" = tidycensus::get_decennial(
      geography = geography,
      variables = variables,
      state = state,
      year = year
    )
  )
  
  # Cache and return
  base::saveRDS(result, cache_file)
  base::message("Cached to: ", cache_file)
  
  result
}
```

```{r geographic-setup}
#| label: geographic-definitions

# Define geographic areas for analysis
geographic_areas_list <- base::list(
  scuc_isd = base::list(
    name = "Schertz-Cibolo-Universal City ISD Area",
    state = "TX",
    # Will use PUMA codes that overlap with school district
    puma_codes = base::c("02104", "02105"), # Approximate PUMA overlap
    description = "Initial test geography - suburban/small city"
  ),
  san_antonio_msa = base::list(
    name = "San Antonio MSA",
    state = "TX", 
    msa_code = "41700",
    description = "Urban comparison geography"
  )
)

base::print("Geographic areas defined for analysis:")
for (area in base::names(geographic_areas_list)) {
  base::cat(base::sprintf("- %s: %s\n", 
              geographic_areas_list[[area]]$name,
              geographic_areas_list[[area]]$description))
}
```

## Variable Definitions

### Success Indicators (Outcome Variables)

```{r success-indicators-data}
#| label: define-success-indicators-data

success_indicators_definitions <- tibble::tribble(
  ~variable_name, ~acs_variable, ~description, ~success_threshold,
  "household_income_median", "HINCP", "Household Income (Past 12 Months)", "> 75th percentile by geography",
  "educational_attainment_bachelor_plus", "SCHL", "Educational Attainment: Bachelor's+", "Code 21-24 (Bachelor's through Doctorate)",
  "employment_status_fulltime", "ESR", "Employment Status", "Code 1 (Employed, at work)",
  "poverty_status_above_threshold", "POVPIP", "Poverty Status (% of poverty line)", "> 200% of poverty threshold",
  "homeownership_status", "TEN", "Housing Tenure", "Code 1-2 (Owned/being bought)",
  "occupation_professional_managerial", "OCCP", "Occupation Categories", "Management/Professional codes"
)
```

::: {.content-visible when-format="html"}
```{r success-indicators-dt}
#| label: success-indicators-dt

DT::datatable(
  success_indicators_definitions,
  caption = "Success Outcome Variables from PUMS Data",
  options = base::list(pageLength = 10, scrollX = TRUE),
  rownames = FALSE
)
```
:::

::: {.content-visible unless-format="html"}
```{r success-indicators-ft}
#| label: success-indicators-ft

flextable::flextable(success_indicators_definitions) |>
  flextable::set_caption(caption = "Success Outcome Variables from PUMS Data") |>
  flextable::add_header_lines(values = "Success Indicators") |>
  flextable::color(i = 1, color = "blue", part = "header") |>
  flextable::italic(i = 1, part = "header") |>
  flextable::align(i = 1, align = "center", part = "header") |>
  flextable::fontsize(i = 1, size = 14, part = "header") |>
  flextable::bg(i = 1, bg = "white", part = "header") |>
  flextable::bg(i = 2, bg = "palegreen", part = "header") |>
  ftExtra::colformat_md() |>
  flextable::autofit()
```
:::

### Sowell Framework Variables (Predictor Variables)

```{r sowell-framework-data}
#| label: define-sowell-framework-data

sowell_framework_variables <- tibble::tribble(
  ~framework_component, ~variable_name, ~acs_variable, ~description, ~coding_logic,
  "High School Graduation", "education_hs_graduate", "SCHL", "Educational Attainment", "Codes 16-24 (HS diploma through doctorate)",
  "Marriage Before Children", "marital_status_at_first_birth", "MSP + FER", "Marriage/Fertility timing", "Married before age at first birth",
  "Employment Maintenance", "employment_consistency", "ESR + WKHP + WKW", "Work history indicators", "Full-time, year-round work pattern",
  "Family Stability Proxy", "household_structure_stable", "HHT + NOC", "Household composition", "Two-parent OR stable single-parent",
  "Educational Values Proxy", "school_enrollment_appropriate", "SCH", "School enrollment patterns", "Age-appropriate enrollment"
)
```

::: {.content-visible when-format="html"}
```{r sowell-framework-dt}
#| label: sowell-framework-dt

DT::datatable(
  sowell_framework_variables,
  caption = "Sowell Framework Variables (Behavioral/Attitudinal Proxies)",
  options = base::list(pageLength = 10, scrollX = TRUE),
  rownames = FALSE
)
```
:::

::: {.content-visible unless-format="html"}
```{r sowell-framework-ft}
#| label: sowell-framework-ft

flextable::flextable(sowell_framework_variables) |>
  flextable::set_caption(caption = "Sowell Framework Variables (Behavioral/Attitudinal Proxies)") |>
  flextable::add_header_lines(values = "Sowell Framework Predictors") |>
  flextable::color(i = 1, color = "blue", part = "header") |>
  flextable::italic(i = 1, part = "header") |>
  flextable::align(i = 1, align = "center", part = "header") |>
  flextable::fontsize(i = 1, size = 14, part = "header") |>
  flextable::bg(i = 1, bg = "white", part = "header") |>
  flextable::bg(i = 2, bg = "palegreen", part = "header") |>
  ftExtra::colformat_md() |>
  flextable::autofit()
```
:::

### Control Variables (Demographic Factors)

```{r control-variables-data}
#| label: define-control-variables-data

demographic_control_variables <- tibble::tribble(
  ~category, ~variable_name, ~acs_variable, ~description,
  "Race/Ethnicity", "race_ethnicity_detailed", "RAC1P + HISP", "Race and Hispanic origin",
  "Economic Status", "family_poverty_status", "POVPIP", "Family poverty level (% of threshold)",
  "Language", "english_proficiency_level", "ENG", "English language proficiency",
  "Family Structure", "family_type_structure", "HHT", "Household type and family structure",
  "Geographic Mobility", "residence_mobility_status", "MIG", "Migration/residential mobility",
  "Age Demographics", "age_category_standardized", "AGEP", "Age in standardized categories"
)
```

::: {.content-visible when-format="html"}
```{r control-variables-dt}
#| label: control-variables-dt

DT::datatable(
  demographic_control_variables,
  caption = "Demographic Control Variables",
  options = base::list(pageLength = 10, scrollX = TRUE),
  rownames = FALSE
)
```
:::

::: {.content-visible unless-format="html"}
```{r control-variables-ft}
#| label: control-variables-ft

flextable::flextable(demographic_control_variables) |>
  flextable::set_caption(caption = "Demographic Control Variables") |>
  flextable::add_header_lines(values = "Demographic Controls") |>
  flextable::color(i = 1, color = "blue", part = "header") |>
  flextable::italic(i = 1, part = "header") |>
  flextable::align(i = 1, align = "center", part = "header") |>
  flextable::fontsize(i = 1, size = 14, part = "header") |>
  flextable::bg(i = 1, bg = "white", part = "header") |>
  flextable::bg(i = 2, bg = "palegreen", part = "header") |>
  ftExtra::colformat_md() |>
  flextable::autofit()
```
:::

## Data Collection Functions

```{r data-collection-functions}
#| label: data-collection-functions

# Function to retrieve PUMS data for specific geography (using cache_census_api_call)
retrieve_pums_data_by_geography <- function(geography_name, year = 2022, cache_dir = "data", force_refresh = FALSE) {
  
  geography_info <- geographic_areas_list[[geography_name]]
  
  base::cat(base::sprintf("Retrieving PUMS data for: %s\n", geography_info$name))
  
  # Define person-level variables
  person_variables <- base::c(
    # Demographics
    "AGEP", "SEX", "RAC1P", "HISP", "ENG",
    # Education
    "SCHL", "SCH", 
    # Employment
    "ESR", "WKHP", "WKW", "OCCP",
    # Income/Poverty
    "POVPIP", "PINCP",
    # Marriage/Family
    "MSP", "FER", "ANC"
  )
  
  # Define household-level variables
  household_variables <- base::c(
    # Housing
    "TEN", "HINCP",
    # Household composition  
    "HHT", "NOC", "NPF"
  )
  
  # Get PUMS person-level data (with caching)
  pums_person_data <- cache_census_api_call(
    type = "pums",
    variables = person_variables,
    state = geography_info$state,
    year = year,
    survey = "acs5",
    cache_dir = cache_dir,
    force_refresh = force_refresh
  )
  
  # Get PUMS household-level data (with caching)
  pums_household_data <- cache_census_api_call(
    type = "pums",
    variables = household_variables,
    state = geography_info$state,
    year = year,
    survey = "acs5",
    cache_dir = cache_dir,
    force_refresh = force_refresh
  )
  
  base::return(base::list(
    person_data = pums_person_data,
    household_data = pums_household_data,
    geography = geography_info$name
  ))
}

# Function to create Sowell Framework compliance score
calculate_sowell_compliance_score <- function(person_data) {
  
  person_data_scored <- person_data |>
    dplyr::mutate(
      # Component 1: High School Graduate (or higher)
      sowell_component_hs_graduate = dplyr::case_when(
        SCHL >= 16 ~ 1,
        SCHL < 16 ~ 0,
        base::is.na(SCHL) ~ NA_real_,
        TRUE ~ NA_real_
      ),
      
      # Component 2: Marriage before children (proxy)
      sowell_component_marriage_first = dplyr::case_when(
        MSP == 1 & (FER == 0 | base::is.na(FER)) ~ 1,
        MSP == 1 & FER > 0 ~ 1,
        MSP != 1 & FER > 0 ~ 0,
        MSP != 1 & (FER == 0 | base::is.na(FER)) ~ NA_real_,
        TRUE ~ NA_real_
      ),
      
      # Component 3: Employment (full-time work)
      sowell_component_employed = dplyr::case_when(
        ESR == 1 & WKHP >= 35 ~ 1,
        ESR == 1 & WKHP < 35 ~ 0.5,
        ESR != 1 ~ 0,
        base::is.na(ESR) ~ NA_real_,
        TRUE ~ NA_real_
      ),
      
      # Calculate overall compliance score (0-3 scale)
      sowell_total_compliance_score = sowell_component_hs_graduate + 
                                    sowell_component_marriage_first + 
                                    sowell_component_employed,
      
      # Create categorical compliance levels
      sowell_compliance_category = dplyr::case_when(
        sowell_total_compliance_score >= 2.5 ~ "High Compliance (2.5-3.0)",
        sowell_total_compliance_score >= 1.5 ~ "Moderate Compliance (1.5-2.4)", 
        sowell_total_compliance_score >= 0.5 ~ "Low Compliance (0.5-1.4)",
        sowell_total_compliance_score >= 0 ~ "Minimal Compliance (0-0.4)",
        base::is.na(sowell_total_compliance_score) ~ "Insufficient Data",
        TRUE ~ "Insufficient Data"
      )
    )
  
  base::return(person_data_scored)
}

# Function to create demographic control categories
categorize_demographic_controls <- function(person_data) {
  
  person_data_categorized <- person_data |>
    dplyr::mutate(
      # Race/ethnicity categories
      race_ethnicity_category = dplyr::case_when(
        HISP != 1 & RAC1P == 1 ~ "Non-Hispanic White",
        HISP != 1 & RAC1P == 2 ~ "Non-Hispanic Black", 
        HISP != 1 & RAC1P == 6 ~ "Non-Hispanic Asian",
        HISP == 1 ~ "Hispanic/Latino",
        TRUE ~ "Other/Multiple Race"
      ),
      
      # Poverty status categories
      poverty_status_category = dplyr::case_when(
        POVPIP < 100 ~ "Below Poverty Line (<100%)",
        POVPIP < 200 ~ "Low Income (100-199%)", 
        POVPIP < 400 ~ "Middle Income (200-399%)",
        POVPIP >= 400 ~ "High Income (400%+)",
        base::is.na(POVPIP) ~ "Income Not Determined"
      ),
      
      # English proficiency categories  
      english_proficiency_category = dplyr::case_when(
        ENG == 1 ~ "English Only",
        ENG == 2 ~ "English Very Well",
        ENG == 3 ~ "English Well", 
        ENG == 4 ~ "English Not Well/Not at All",
        base::is.na(ENG) ~ "Not Applicable/Unknown"
      ),
      
      # Age categories for analysis
      age_analysis_category = dplyr::case_when(
        AGEP < 18 ~ "Under 18 (Minors)",
        AGEP >= 18 & AGEP < 25 ~ "Young Adults (18-24)",
        AGEP >= 25 & AGEP < 35 ~ "Adults (25-34)",
        AGEP >= 35 & AGEP < 50 ~ "Middle Age (35-49)", 
        AGEP >= 50 & AGEP < 65 ~ "Older Adults (50-64)",
        AGEP >= 65 ~ "Seniors (65+)",
        TRUE ~ "Age Unknown"
      )
    )
  
  base::return(person_data_categorized)
}
```

## Phase 1 Analysis: Schertz-Cibolo-Universal City Area

```{r phase1-data-collection}
#| label: phase1-data-collection
#| eval: false

# Note: This chunk is set to eval=false because it requires a Census API key

# Collect Phase 1 data
phase1_raw_data <- retrieve_pums_data_by_geography("scuc_isd", year = 2022)

# Process the person-level data
phase1_person_analysis <- phase1_raw_data$person_data |>
  calculate_sowell_compliance_score() |>
  categorize_demographic_controls()

# Create summary statistics
phase1_summary_by_compliance <- phase1_person_analysis |>
  dplyr::filter(!base::is.na(sowell_total_compliance_score)) |>
  dplyr::group_by(sowell_compliance_category, race_ethnicity_category, poverty_status_category) |>
  dplyr::summarise(
    sample_size = dplyr::n(),
    median_personal_income = stats::median(PINCP, na.rm = TRUE),
    pct_bachelor_plus = base::mean(SCHL >= 21, na.rm = TRUE) * 100,
    pct_employed_fulltime = base::mean(ESR == 1 & WKHP >= 35, na.rm = TRUE) * 100,
    .groups = "drop"
  )
```

## Sample Analysis Framework (Using Simulated Data)

Since running the full analysis requires a Census API key, here's the analytical framework using simulated data that matches the expected structure:

```{r simulated-analysis-data}
#| label: simulated-analysis-data

# Create simulated data for demonstration
base::set.seed(12345)
n_sample <- 5000

simulated_census_data <- tibble::tibble(
  # Sowell Framework Components
  sowell_component_hs_graduate = base::sample(
    base::c(0, 1), n_sample, replace = TRUE, prob = base::c(0.15, 0.85)
  ),
  sowell_component_marriage_first = base::sample(
    base::c(0, 1), n_sample, replace = TRUE, prob = base::c(0.35, 0.65)
  ),
  sowell_component_employed = base::sample(
    base::c(0, 0.5, 1), n_sample, replace = TRUE, prob = base::c(0.15, 0.15, 0.70)
  ),
  
  # Demographics
  race_ethnicity_category = base::sample(
    base::c("Non-Hispanic White", "Non-Hispanic Black", "Hispanic/Latino", 
            "Non-Hispanic Asian", "Other/Multiple Race"),
    n_sample, replace = TRUE, prob = base::c(0.45, 0.15, 0.25, 0.10, 0.05)
  ),
  poverty_status_category = base::sample(
    base::c("Below Poverty Line (<100%)", "Low Income (100-199%)", 
            "Middle Income (200-399%)", "High Income (400%+)"),
    n_sample, replace = TRUE, prob = base::c(0.15, 0.25, 0.35, 0.25)
  ),
  english_proficiency_category = base::sample(
    base::c("English Only", "English Very Well", "English Well", 
            "English Not Well/Not at All"),
    n_sample, replace = TRUE, prob = base::c(0.70, 0.15, 0.10, 0.05)
  )
) |>
  dplyr::mutate(
    sowell_total_compliance_score = sowell_component_hs_graduate + 
                                  sowell_component_marriage_first + 
                                  sowell_component_employed,
    
    sowell_compliance_category = dplyr::case_when(
      sowell_total_compliance_score >= 2.5 ~ "High Compliance (2.5-3.0)",
      sowell_total_compliance_score >= 1.5 ~ "Moderate Compliance (1.5-2.4)", 
      sowell_total_compliance_score >= 0.5 ~ "Low Compliance (0.5-1.4)",
      TRUE ~ "Minimal Compliance (0-0.4)"
    ),
    
    success_probability = 0.15 + (sowell_total_compliance_score * 0.25) + 
                         stats::rnorm(n_sample, 0, 0.1),
    success_probability = base::pmax(0, base::pmin(1, success_probability)),
    
    personal_income = base::exp(9 + (sowell_total_compliance_score * 0.4) + 
                         stats::rnorm(n_sample, 0, 0.3)) * 
                     base::ifelse(poverty_status_category == "High Income (400%+)", 1.3, 1),
    bachelor_plus = stats::rbinom(n_sample, 1, success_probability),
    homeownership = stats::rbinom(n_sample, 1, success_probability * 0.8)
  )

# Analysis 1: Success outcomes by Sowell compliance across demographic groups
success_by_compliance_demographics <- simulated_census_data |>
  dplyr::group_by(sowell_compliance_category, race_ethnicity_category, poverty_status_category) |>
  dplyr::summarise(
    sample_size = dplyr::n(),
    median_income = stats::median(personal_income, na.rm = TRUE),
    pct_bachelor_plus = base::mean(bachelor_plus, na.rm = TRUE) * 100,
    pct_homeowner = base::mean(homeownership, na.rm = TRUE) * 100,
    mean_compliance_score = base::mean(sowell_total_compliance_score, na.rm = TRUE),
    .groups = "drop"
  ) |>
  dplyr::filter(sample_size >= 30)
```

::: {.content-visible when-format="html"}
```{r simulated-analysis-dt}
#| label: simulated-analysis-dt

DT::datatable(
  utils::head(success_by_compliance_demographics, 20),
  caption = "Success Outcomes by Sowell Compliance and Demographics (Simulated Data)",
  options = base::list(pageLength = 15, scrollX = TRUE, dom = 'Bfrtip'),
  rownames = FALSE
) |>
  DT::formatRound(
    columns = base::c("median_income", "pct_bachelor_plus", "pct_homeowner", "mean_compliance_score"), 
    digits = 2
  )
```
:::

::: {.content-visible unless-format="html"}
```{r simulated-analysis-ft}
#| label: simulated-analysis-ft

flextable::flextable(utils::head(success_by_compliance_demographics, 20)) |>
  flextable::set_caption(caption = "Success Outcomes by Sowell Compliance and Demographics (Simulated Data)") |>
  flextable::add_header_lines(values = "Success by Compliance Level") |>
  flextable::color(i = 1, color = "blue", part = "header") |>
  flextable::italic(i = 1, part = "header") |>
  flextable::align(i = 1, align = "center", part = "header") |>
  flextable::fontsize(i = 1, size = 14, part = "header") |>
  flextable::bg(i = 1, bg = "white", part = "header") |>
  flextable::bg(i = 2, bg = "palegreen", part = "header") |>
  flextable::colformat_double(j = c("median_income", "pct_bachelor_plus", "pct_homeowner", "mean_compliance_score"), digits = 2) |>
  ftExtra::colformat_md() |>
  flextable::autofit()
```
:::

```{r key-hypothesis-test-data}
#| label: key-hypothesis-test-data

# Test the core hypothesis: Do high-compliance individuals succeed regardless of demographics?
hypothesis_test_results <- simulated_census_data |>
  dplyr::filter(sowell_compliance_category == "High Compliance (2.5-3.0)") |>
  dplyr::group_by(race_ethnicity_category, poverty_status_category) |>
  dplyr::summarise(
    sample_size = dplyr::n(),
    median_income = stats::median(personal_income, na.rm = TRUE),
    income_75th_percentile = stats::quantile(personal_income, 0.75, na.rm = TRUE),
    pct_bachelor_plus = base::mean(bachelor_plus, na.rm = TRUE) * 100,
    pct_homeowner = base::mean(homeownership, na.rm = TRUE) * 100,
    success_rate_composite = (pct_bachelor_plus * 0.4 + pct_homeowner * 0.3 + 
                             (median_income > 50000) * 30),
    .groups = "drop"
  ) |>
  dplyr::arrange(dplyr::desc(success_rate_composite))
```

::: {.content-visible when-format="html"}
```{r key-hypothesis-test-dt}
#| label: key-hypothesis-test-dt

DT::datatable(
  hypothesis_test_results,
  caption = "HIGH COMPLIANCE Group: Success Rates Across All Demographics",
  options = base::list(pageLength = 15, scrollX = TRUE, dom = 'Bfrtip'),
  rownames = FALSE
) |>
  DT::formatRound(
    columns = base::c("median_income", "income_75th_percentile", "pct_bachelor_plus", 
                     "pct_homeowner", "success_rate_composite"), 
    digits = 2
  )
```
:::

::: {.content-visible unless-format="html"}
```{r key-hypothesis-test-ft}
#| label: key-hypothesis-test-ft

flextable::flextable(hypothesis_test_results) |>
  flextable::set_caption(caption = "HIGH COMPLIANCE Group: Success Rates Across All Demographics") |>
  flextable::add_header_lines(values = "High Compliance Success Analysis") |>
  flextable::color(i = 1, color = "blue", part = "header") |>
  flextable::italic(i = 1, part = "header") |>
  flextable::align(i = 1, align = "center", part = "header") |>
  flextable::fontsize(i = 1, size = 14, part = "header") |>
  flextable::bg(i = 1, bg = "white", part = "header") |>
  flextable::bg(i = 2, bg = "palegreen", part = "header") |>
  flextable::colformat_double(j = c("median_income", "income_75th_percentile", "pct_bachelor_plus", 
                                    "pct_homeowner", "success_rate_composite"), digits = 2) |>
  ftExtra::colformat_md() |>
  flextable::autofit()
```
:::

```{r hypothesis-variance}
#| label: hypothesis-variance

# Calculate the variance in success rates within high compliance group
success_variance_high_compliance <- stats::var(hypothesis_test_results$success_rate_composite, na.rm = TRUE)

base::cat(base::sprintf("\nVariance in success rates among HIGH COMPLIANCE individuals: %.2f\n", 
            success_variance_high_compliance))
base::cat("Lower variance suggests demographic factors matter less for high-compliance individuals.\n")
```

## Visualization Framework

```{r visualization-framework}
#| label: create-visualizations

# Visualization 1: Success rates by compliance level, faceted by demographics
success_by_compliance_plot <- simulated_census_data |>
  dplyr::group_by(sowell_compliance_category, race_ethnicity_category) |>
  dplyr::summarise(
    median_income = stats::median(personal_income, na.rm = TRUE),
    pct_bachelor = base::mean(bachelor_plus, na.rm = TRUE) * 100,
    sample_size = dplyr::n(),
    .groups = "drop"
  ) |>
  dplyr::filter(sample_size >= 50) |>
  ggplot2::ggplot(ggplot2::aes(x = sowell_compliance_category, y = pct_bachelor, 
                              fill = race_ethnicity_category)) +
  ggplot2::geom_col(position = "dodge") +
  ggplot2::facet_wrap(~race_ethnicity_category, scales = "free_y") +
  ggplot2::labs(
    title = "Bachelor's Degree Attainment by Sowell Compliance Level",
    subtitle = "Across Different Racial/Ethnic Groups", 
    x = "Sowell Framework Compliance Level",
    y = "Percentage with Bachelor's Degree or Higher",
    fill = "Race/Ethnicity"
  )

base::print(success_by_compliance_plot)

# Visualization 2: Income distribution by compliance, controlling for poverty status
income_distribution_plot <- simulated_census_data |>
  ggplot2::ggplot(ggplot2::aes(x = sowell_compliance_category, y = personal_income,
                              fill = poverty_status_category)) +
  ggplot2::geom_boxplot() +
  ggplot2::scale_y_log10(labels = scales::dollar) +
  ggplot2::facet_wrap(~poverty_status_category, scales = "free_y") +
  ggplot2::labs(
    title = "Personal Income Distribution by Sowell Compliance",
    subtitle = "Controlling for Initial Poverty Status",
    x = "Sowell Framework Compliance Level", 
    y = "Personal Income (Log Scale)",
    fill = "Poverty Status"
  )

base::print(income_distribution_plot)
```

## Statistical Testing Framework

```{r statistical-tests-data}
#| label: statistical-testing-data

# Test 1: Does Sowell compliance predict success better than demographics alone?

# Model 1: Demographics only
model_demographics_only <- stats::lm(
  personal_income ~ race_ethnicity_category + poverty_status_category + 
                   english_proficiency_category,
  data = simulated_census_data
)

# Model 2: Sowell compliance only  
model_sowell_only <- stats::lm(
  personal_income ~ sowell_total_compliance_score,
  data = simulated_census_data
)

# Model 3: Combined model
model_combined <- stats::lm(
  personal_income ~ sowell_total_compliance_score + race_ethnicity_category + 
                   poverty_status_category + english_proficiency_category,
  data = simulated_census_data
)

# Extract R-squared values
r_squared_demographics <- base::summary(model_demographics_only)$r.squared
r_squared_sowell <- base::summary(model_sowell_only)$r.squared  
r_squared_combined <- base::summary(model_combined)$r.squared

model_comparison_results <- tibble::tibble(
  Model = base::c("Demographics Only", "Sowell Framework Only", "Combined Model"),
  R_Squared = base::c(r_squared_demographics, r_squared_sowell, r_squared_combined),
  Adjusted_R_Squared = base::c(
    base::summary(model_demographics_only)$adj.r.squared,
    base::summary(model_sowell_only)$adj.r.squared,
    base::summary(model_combined)$adj.r.squared
  )
)
```

::: {.content-visible when-format="html"}
```{r statistical-tests-dt}
#| label: statistical-tests-dt

DT::datatable(
  model_comparison_results,
  caption = "Model Comparison: Predictive Power of Different Frameworks",
  options = base::list(pageLength = 10, scrollX = TRUE),
  rownames = FALSE
) |>
  DT::formatRound(columns = base::c("R_Squared", "Adjusted_R_Squared"), digits = 4)
```
:::

::: {.content-visible unless-format="html"}
```{r statistical-tests-ft}
#| label: statistical-tests-ft

flextable::flextable(model_comparison_results) |>
  flextable::set_caption(caption = "Model Comparison: Predictive Power of Different Frameworks") |>
  flextable::add_header_lines(values = "Model Comparison Results") |>
  flextable::color(i = 1, color = "blue", part = "header") |>
  flextable::italic(i = 1, part = "header") |>
  flextable::align(i = 1, align = "center", part = "header") |>
  flextable::fontsize(i = 1, size = 14, part = "header") |>
  flextable::bg(i = 1, bg = "white", part = "header") |>
  flextable::bg(i = 2, bg = "palegreen", part = "header") |>
  flextable::colformat_double(j = c("R_Squared", "Adjusted_R_Squared"), digits = 4) |>
  ftExtra::colformat_md() |>
  flextable::autofit()
```
:::

```{r interaction-model}
#| label: interaction-model

# Test 2: Interaction effects - does Sowell compliance work differently across groups?
interaction_model <- stats::lm(
  personal_income ~ sowell_total_compliance_score * race_ethnicity_category * poverty_status_category,
  data = simulated_census_data
)

base::cat("\nInteraction Model Summary (showing if Sowell framework works differently across groups):\n")
base::print(base::summary(interaction_model)$coefficients[1:10, ])
```

## Implementation Roadmap

### Phase 1: Current Geographic Focus
1. **Data Collection**: Schertz-Cibolo-Universal City ISD area PUMS data
2. **Baseline Analysis**: Establish Sowell compliance rates and success metrics
3. **Initial Hypothesis Testing**: Does framework predict success across demographics?

### Phase 2: Geographic Expansion  
1. **San Antonio MSA Analysis**: Test in urban environment
2. **Comparative Analysis**: Urban vs suburban patterns
3. **Robustness Testing**: Do results hold across geographic contexts?

### Phase 3: Variable Iteration
1. **One-Variable Changes**: Systematically alter individual demographic controls
2. **Sensitivity Analysis**: Where does the hypothesis break down?
3. **Refinement**: Identify most critical framework components

### Limitations and Considerations

#### Data Limitations
- **Attitude/Values Proxies**: Census data cannot directly measure family values or attitudes
- **Causation vs Correlation**: Analysis shows associations, not causal relationships
- **Temporal Sequencing**: Difficult to establish timing of marriage/children/education
- **Selection Effects**: High-compliance individuals may differ in unobservable ways

#### Methodological Considerations
- **Sample Size Requirements**: Need adequate representation across demographic groups
- **Geographic Specificity**: Results may vary significantly by local context
- **Economic Cycle Effects**: Labor market conditions affect employment outcomes
- **Measurement Error**: Self-reported data and proxy variables introduce uncertainty

#### Ethical Research Notes
- **Avoid Victim Blaming**: Results should inform policy, not blame individuals
- **Structural Factors**: Individual choices occur within broader structural constraints
- **Policy Implications**: Focus on enabling success rather than mandating behaviors

## Next Steps for Implementation

```{r next-steps-checklist-data}
#| label: implementation-checklist-data

next_steps_checklist <- tibble::tribble(
  ~Step, ~Action_Required, ~Priority, ~Dependencies,
  "1", "Obtain Census API key from census.gov", "High", "None",
  "2", "Validate geographic PUMA codes for target areas", "High", "API access", 
  "3", "Run Phase 1 data collection for SCUC ISD", "High", "Steps 1-2",
  "4", "Refine variable definitions based on actual data availability", "Medium", "Step 3",
  "5", "Implement statistical testing framework", "Medium", "Steps 3-4",
  "6", "Expand to San Antonio MSA (Phase 2)", "Medium", "Phase 1 completion",
  "7", "Create automated reporting pipeline", "Low", "Phases 1-2 validation",
  "8", "Plan additional geographic areas for Phase 3", "Low", "Phases 1-2 success"
)
```

::: {.content-visible when-format="html"}
```{r next-steps-checklist-dt}
#| label: next-steps-checklist-dt

DT::datatable(
  next_steps_checklist,
  caption = "Implementation Checklist for Full Analysis",
  options = base::list(pageLength = 10, scrollX = TRUE),
  rownames = FALSE
)
```
:::

::: {.content-visible unless-format="html"}
```{r next-steps-checklist-ft}
#| label: next-steps-checklist-ft

flextable::flextable(next_steps_checklist) |>
  flextable::set_caption(caption = "Implementation Checklist for Full Analysis") |>
  flextable::add_header_lines(values = "Implementation Checklist") |>
  flextable::color(i = 1, color = "blue", part = "header") |>
  flextable::italic(i = 1, part = "header") |>
  flextable::align(i = 1, align = "center", part = "header") |>
  flextable::fontsize(i = 1, size = 14, part = "header") |>
  flextable::bg(i = 1, bg = "white", part = "header") |>
  flextable::bg(i = 2, bg = "palegreen", part = "header") |>
  ftExtra::colformat_md() |>
  flextable::autofit()
```
:::

---

**Research Framework Summary**: This analysis framework provides a rigorous approach to testing whether Thomas Sowell's success framework (high school graduation + marriage before children + employment) predicts positive outcomes across demographic lines. The methodology uses individual-level Census data to control for race, poverty, language, and family structure while measuring behavioral compliance with success-associated behaviors.

The key innovation is using Census microdata to create behavioral compliance scores that serve as proxies for family values/attitudes, then testing whether these predict success better than demographic characteristics alone.
