---
title: "Enhanced Census Longitudinal Analysis with Shiny Integration"
subtitle: "Cibolo vs Schertz Analysis with Optional Data Standardization"
description: "Flexible longitudinal analysis with optional Shiny app data formatting"
author: "Census Analysis Tool"
date: today
format:
  html:
    code-copy: true
    page-layout: full
    fig-width: 12
    fig-height: 10
    fig-dpi: 300
    df-print: paged
    code-overflow: wrap
    toc: true
    toc-depth: 3
  typst:
    fig-width: 12
    fig-height: 10
    fig-dpi: 300
    margin:
      x: 1in
      y: 1in
    toc: true
    fontsize: 14pt
    mainfont: "Latin Modern Roman"
execute:
  echo: true
  message: false
  warning: false
  eval: true
  fig-width: 12
  fig-height: 10
---

```{r setup}
#| label: setup
#| include: false

knitr::opts_chunk$set(echo = TRUE)

# Prevent scientific notation globally
options(scipen = 999)

# Set seed for reproducibility
set.seed(123)
```


> **Note:** Run chunks sequentially the first time. After initial run, 
> you can modify parameters and re-run specific sections as needed.


## Enhanced Longitudinal Census Analysis

This analysis provides flexible longitudinal data collection and visualization for US Census data, with optional standardization for Shiny dashboard integration.

### Key Features:
- **14 years maximum data range** (2009-2022)
- **User-defined analysis years** at the top
- **Optional Shiny standardization** with toggle parameter
- **Professional line charts** with trend analysis
- **Multiple export formats** (.csv, .xlsx, .Rds)

---

```{r main-analysis}
# Enhanced Longitudinal Census Analysis: Cibolo vs Schertz
# With Optional Shiny App Standardization

# Load required libraries
library(tidycensus)
library(tidyverse)
library(writexl)
library(scales)
library(ggplot2)

```


#==============================================================================
# STEP 1: SET YOUR ANALYSIS PARAMETERS (USER CUSTOMIZATION)
#==============================================================================

```{r}

# SET YOUR YEARS HERE - Customize based on how far back you want to go
ANALYSIS_YEARS <- 2009:2022  # Maximum ACS 5-year range (14 years!)
# Alternative options:
# ANALYSIS_YEARS <- 2015:2022  # Recent 8 years (faster processing)
# ANALYSIS_YEARS <- c(2009, 2012, 2015, 2018, 2022)  # Every 3 years
# ANALYSIS_YEARS <- 2010:2022  # Post-recession era
# ANALYSIS_YEARS <- 2013:2019  # Pre-pandemic era

# Set your cities
CITIES <- c("Cibolo city, Texas", "Schertz city, Texas")

# Choose your survey type
SURVEY_TYPE <- "acs5"  # Use "acs1" if cities were >65k (they're not)

# SHINY APP INTEGRATION TOGGLE
STANDARDIZE_FOR_SHINY <- TRUE  # Set to TRUE to create Shiny-ready datasets

```

#==============================================================================
# DATA AVAILABILITY REFERENCE
#==============================================================================

```{r}

cat("=== ENHANCED LONGITUDINAL CENSUS ANALYSIS ===\n")
cat("ACS 5-Year Estimates: 2009-2022 (14 years maximum)\n")
cat("Decennial Census: 1990, 2000, 2010, 2020 (limited variables)\n")
cat("Population Estimates: 2000-2022 (population only)\n\n")

cat("SELECTED ANALYSIS YEARS:", paste(ANALYSIS_YEARS, collapse = ", "), "\n")
cat("TOTAL YEARS OF DATA:", length(ANALYSIS_YEARS), "\n")
cat("ANALYSIS SPAN:", max(ANALYSIS_YEARS) - min(ANALYSIS_YEARS) + 1, "years\n")
cat("SHINY STANDARDIZATION:", ifelse(STANDARDIZE_FOR_SHINY, "ENABLED", "DISABLED"), "\n\n")

```

#==============================================================================
# STEP 2: DEFINE VARIABLES FOR ANALYSIS
#==============================================================================

```{r}

# Variables optimized for meaningful trend visualization
longitudinal_vars <- c(
  # Population and Demographics
  total_pop = "DP05_0001",
  median_age = "DP05_0018",
  under_18_pct = "DP05_0019P",
  over_65_pct = "DP05_0024P",
  hispanic_pct = "DP05_0071P",
  
  # Education and Social
  bachelor_plus_pct = "DP02_0065P",
  married_couple_families_pct = "DP02_0007P",
  avg_household_size = "DP02_0016",
  
  # Economic Indicators
  median_hh_income = "DP03_0062",
  per_capita_income = "DP03_0088", 
  unemployment_rate = "DP03_0005P",
  poverty_rate = "DP03_0119P",
  
  # Housing Market
  median_home_value = "DP04_0089",
  median_gross_rent = "DP04_0134",
  owner_occupied_pct = "DP04_0046P",
  vacancy_rate = "DP04_0003P"
)

```

#==============================================================================
# STEP 3: DATA COLLECTION FUNCTIONS
#==============================================================================

```{r}

# Function to check data availability
check_chart_data_availability <- function(years, cities, variables) {
  cat("=== CHECKING DATA AVAILABILITY ===\n")
  cat("Testing years:", paste(range(years), collapse = " to "), "\n")
  cat("Cities:", paste(cities, collapse = ", "), "\n\n")
  
  # Test first, middle, and last year
  test_years <- unique(c(min(years), median(years), max(years)))
  
  for(yr in test_years) {
    cat("Testing year", yr, "... ")
    
    tryCatch({
      test_data <- get_acs(
        geography = "place",
        variables = variables[1:3], # Test with just a few variables
        state = "TX",
        year = yr,
        survey = SURVEY_TYPE
      ) %>%
        filter(NAME %in% cities)
      
      if(nrow(test_data) > 0) {
        cat("✓ Available (", nrow(test_data), "records)\n")
      } else {
        cat("✗ No data for these cities\n")
      }
    }, error = function(e) {
      cat("✗ Error:", e$message, "\n")
    })
  }
  cat("\n")
}

# Enhanced longitudinal data function
get_longitudinal_data <- function(years = ANALYSIS_YEARS,
                                 cities = CITIES, 
                                 variables = longitudinal_vars,
                                 survey = SURVEY_TYPE) {
  
  cat("=== STARTING LONGITUDINAL DATA COLLECTION ===\n")
  cat("Years:", paste(years, collapse = ", "), "\n")
  cat("Variables:", length(variables), "\n")
  cat("Survey:", survey, "\n\n")
  
  # Get data for all years and cities
  longitudinal_data <- map_dfr(years, function(yr) {
    
    cat("Fetching data for year:", yr, "...")
    
    tryCatch({
      yearly_data <- get_acs(
        geography = "place",
        variables = variables,
        state = "TX",
        year = yr,
        survey = survey
      ) %>%
        filter(NAME %in% cities) %>%
        mutate(
          year = yr,
          city = case_when(
            str_detect(NAME, "Cibolo") ~ "Cibolo",
            str_detect(NAME, "Schertz") ~ "Schertz",
            TRUE ~ NAME
          ),
          # Create period labels for ACS 5-year estimates
          period = if(survey == "acs5") paste0(yr-4, "-", yr) else as.character(yr)
        ) %>%
        select(city, year, period, variable, estimate, moe)
      
      cat(" ✓ Success (", nrow(yearly_data), " records)\n")
      return(yearly_data)
      
    }, error = function(e) {
      cat(" ✗ Error:", e$message, "\n")
      return(tibble())
    })
  })
  
  if(nrow(longitudinal_data) == 0) {
    stop("No data retrieved. Check your years, cities, and variables.")
  }
  
  # Add variable labels and categories
  longitudinal_data <- longitudinal_data %>%
    mutate(
      variable_label = case_when(
        variable == "total_pop" ~ "Total Population",
        variable == "median_age" ~ "Median Age",
        variable == "under_18_pct" ~ "Percent Under 18",
        variable == "over_65_pct" ~ "Percent Over 65", 
        variable == "hispanic_pct" ~ "Percent Hispanic/Latino",
        variable == "bachelor_plus_pct" ~ "Percent Bachelor's Degree+",
        variable == "married_couple_families_pct" ~ "Percent Married Couple Families",
        variable == "avg_household_size" ~ "Average Household Size",
        variable == "median_hh_income" ~ "Median Household Income",
        variable == "per_capita_income" ~ "Per Capita Income",
        variable == "unemployment_rate" ~ "Unemployment Rate (%)",
        variable == "poverty_rate" ~ "Poverty Rate (%)",
        variable == "median_home_value" ~ "Median Home Value",
        variable == "median_gross_rent" ~ "Median Gross Rent", 
        variable == "owner_occupied_pct" ~ "Percent Owner Occupied",
        variable == "vacancy_rate" ~ "Vacancy Rate (%)",
        TRUE ~ variable
      ),
      # Add categories for grouping
      category = case_when(
        variable %in% c("total_pop", "median_age", "under_18_pct", "over_65_pct", "hispanic_pct") ~ "Demographics",
        variable %in% c("bachelor_plus_pct", "married_couple_families_pct", "avg_household_size") ~ "Social Characteristics",
        variable %in% c("median_hh_income", "per_capita_income", "unemployment_rate", "poverty_rate") ~ "Economic",
        variable %in% c("median_home_value", "median_gross_rent", "owner_occupied_pct", "vacancy_rate") ~ "Housing",
        TRUE ~ "Other"
      )
    )
  
  cat("\nData collection complete!\n")
  cat("Total records:", nrow(longitudinal_data), "\n")
  cat("Years with data:", paste(sort(unique(longitudinal_data$year)), collapse = ", "), "\n")
  cat("Variables collected:", length(unique(longitudinal_data$variable)), "\n\n")
  
  return(longitudinal_data)
}

```


#==============================================================================
# STEP 4: SHINY STANDARDIZATION FUNCTIONS
#==============================================================================

```{r}

# Function to standardize data for Shiny app
standardize_for_shiny_app <- function(data) {
  cat("=== STANDARDIZING DATA FOR SHINY APP ===\n")
  
  # Transform data to Shiny format: date, value, [categories...]
  shiny_data <- data %>%
    rename(
      date = year,        # year becomes date (x-axis for Shiny)
      value = estimate    # estimate becomes value (y-axis for Shiny)
    ) %>%
    # Reorder columns: date, value, then all categorical columns
    select(date, value, city, period, category, variable_label, moe, variable) %>%
    arrange(date, city, category, variable_label)
  
  cat("Shiny standardization complete!\n")
  cat("Structure: date, value, + ", ncol(shiny_data) - 2, " categorical columns\n")
  cat("Total records:", nrow(shiny_data), "\n\n")
  
  return(shiny_data)
}

```


```{r}

# Function to save standardized data in multiple formats
save_shiny_data <- function(shiny_data, base_filename) {
  year_range <- paste0(min(ANALYSIS_YEARS), "_", max(ANALYSIS_YEARS))

  # Create filenames
  csv_filename <- paste0(base_filename, "_shiny_", year_range, ".csv")
  rds_filename <- paste0(base_filename, "_shiny_", year_range, ".rds")

#   # Save as CSV (universal access)
#   write_csv(shiny_data, csv_filename)
#   cat("✓ Shiny data saved as CSV:", csv_filename, "\n")
#   
#   # Save as RDS (R native format, preserves data types)
#   saveRDS(shiny_data, rds_filename)
#   cat("✓ Shiny data saved as RDS:", rds_filename, "\n")
#   
  # Display summary for Shiny app usage
  cat("\n=== SHINY APP INTEGRATION SUMMARY ===\n")
  cat("Dataset structure for Shiny:\n")
  cat("- First column 'date':", class(shiny_data$date), "(time axis)\n")
  cat("- Second column 'value':", class(shiny_data$value), "(measurement axis)\n")
  cat("- Categorical columns:", paste(names(shiny_data)[3:ncol(shiny_data)], collapse = ", "), "\n")
  cat("- Total combinations:", length(unique(paste(shiny_data$city, shiny_data$variable_label))), "\n")
  cat("- Years covered:", paste(sort(unique(shiny_data$date)), collapse = ", "), "\n\n")

  return(list(csv = csv_filename, rds = rds_filename))
}

```


#==============================================================================
# STEP 5: LINE CHART FUNCTIONS
#==============================================================================

```{r}

# Enhanced function to create line charts for any variable
create_line_chart <- function(data, var_name, title_suffix = "", years_in_title = TRUE) {
  
  chart_data <- data %>%
    filter(variable_label == var_name)
  
  if(nrow(chart_data) == 0) {
    cat("Variable '", var_name, "' not found.\n")
    cat("Available variables:\n")
    print(unique(data$variable_label))
    return(NULL)
  }
  
  # Determine formatting based on variable type
  is_dollar <- str_detect(var_name, "Income|Value|Rent")
  is_percent <- str_detect(var_name, "Percent|Rate|%")
  
  # Create title with year range if requested
  year_range <- if(years_in_title) {
    paste0(" (", min(chart_data$year), "-", max(chart_data$year), ")")
  } else {
    ""
  }
  
  p <- ggplot(chart_data, aes(x = year, y = estimate, color = city, group = city)) +
    geom_line(size = 1.2) +
    geom_point(size = 3) +
    labs(
      title = paste0(var_name, " Comparison: Cibolo vs Schertz", year_range, title_suffix),
      subtitle = paste0("ACS 5-Year Estimates, ", length(unique(chart_data$year)), " years of data"),
      x = "Year",
      y = var_name,
      color = "City",
      caption = "Source: US Census Bureau, American Community Survey"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold"),
      plot.subtitle = element_text(size = 12),
      legend.position = "bottom",
      panel.grid.minor = element_blank(),
      axis.text.x = element_text(angle = 45, hjust = 1)
    ) +
    scale_color_manual(values = c("Cibolo" = "#1f77b4", "Schertz" = "#ff7f0e")) +
    scale_x_continuous(breaks = unique(chart_data$year))
  
  # Add appropriate y-axis formatting
  if(is_dollar) {
    p <- p + scale_y_continuous(labels = dollar_format())
  } else if(is_percent) {
    p <- p + scale_y_continuous(labels = function(x) paste0(x, "%"))
  } else if(max(chart_data$estimate, na.rm = TRUE) > 1000) {
    p <- p + scale_y_continuous(labels = comma_format())
  }
  
  return(p)
}

```

```{r}

# Function to create and save multiple charts at once
create_chart_dashboard <- function(data, variables_to_plot, save_plots = TRUE) {
  
  cat("=== CREATING CHART DASHBOARD ===\n")
  
  charts <- map(variables_to_plot, function(var) {
    cat("Creating chart for:", var, "\n")
    chart <- create_line_chart(data, var)
    
    if(save_plots && !is.null(chart)) {
      year_range <- paste0(min(ANALYSIS_YEARS), "_", max(ANALYSIS_YEARS))
      filename <- paste0(str_replace_all(str_to_lower(var), "[^a-z0-9]", "_"), 
                        "_", year_range, ".png")
      ggsave(filename, chart, width = 12, height = 8, dpi = 300)
      cat("  → Saved as:", filename, "\n")
    }
    
    return(chart)
  })
  
  names(charts) <- variables_to_plot
  cat("Dashboard complete! Created", length(charts), "charts.\n\n")
  
  return(charts)
}

```

```{r}

# Function to analyze trends for any variable
analyze_variable_trends <- function(data, var_name) {
  
  trend_data <- data %>%
    filter(variable_label == var_name) %>%
    group_by(city) %>%
    arrange(year) %>%
    summarise(
      years_available = n(),
      first_year = min(year),
      last_year = max(year),
      first_value = first(estimate),
      last_value = last(estimate),
      total_change = last_value - first_value,
      total_growth_pct = round(((last_value / first_value) - 1) * 100, 1),
      cagr = round((((last_value / first_value)^(1/(last_year - first_year))) - 1) * 100, 2),
      .groups = "drop"
    )
  
  # Year-by-year comparison
  yearly_comparison <- data %>%
    filter(variable_label == var_name) %>%
    select(city, year, estimate) %>%
    pivot_wider(names_from = city, values_from = estimate) %>%
    mutate(
      difference = Schertz - Cibolo,
      leader = ifelse(Schertz > Cibolo, "Schertz", "Cibolo"),
      gap_pct = round(abs(difference) / pmin(Schertz, Cibolo) * 100, 1)
    )
  
  cat("=== TREND ANALYSIS:", var_name, "===\n")
  cat("Growth Summary:\n")
  print(trend_data)
  cat("\nYear-by-Year Comparison:\n") 
  print(yearly_comparison)
  cat("\n")
  
  return(list(growth = trend_data, yearly = yearly_comparison))
}

```


#==============================================================================
# STEP 6: RUN THE ANALYSIS
#==============================================================================

```{r}

# Check data availability
check_chart_data_availability(ANALYSIS_YEARS, CITIES, longitudinal_vars)

# Get the longitudinal data
longitudinal_data <- get_longitudinal_data()

# Create separate datasets for each city
cibolo_data <- longitudinal_data %>%
  filter(city == "Cibolo") %>%
  arrange(variable, year)

schertz_data <- longitudinal_data %>%
  filter(city == "Schertz") %>%
  arrange(variable, year)

# Create wide format for Excel analysis
longitudinal_wide <- longitudinal_data %>%
  select(city, year, variable_label, estimate) %>%
  pivot_wider(
    names_from = city,
    values_from = estimate,
    names_prefix = ""
  ) %>%
  group_by(variable_label) %>%
  mutate(
    difference = Schertz - Cibolo,
    schertz_growth_rate = round((Schertz / lag(Schertz, order_by = year) - 1) * 100, 1),
    cibolo_growth_rate = round((Cibolo / lag(Cibolo, order_by = year) - 1) * 100, 1)
  ) %>%
  ungroup() %>%
  arrange(variable_label, year)

```

#==============================================================================
# STEP 7: SHINY STANDARDIZATION (CONDITIONAL)
#==============================================================================

```{r}

# if(STANDARDIZE_FOR_SHINY) {
#   # Standardize data for Shiny app
#   shiny_standardized_data <- standardize_for_shiny_app(longitudinal_data)
#   
#   # Save in both formats
#   shiny_files <- save_shiny_data(shiny_standardized_data, "cibolo_schertz_census")
#   
#   cat("=== SHINY APP USAGE INSTRUCTIONS ===\n")
#   cat("1. Upload either the .csv or .rds file to your Shiny app\n")
#   cat("2. Data structure: 'date' (time), 'value' (measurement), + categories\n")
#   cat("3. Filter by 'variable_label' to select indicators\n")
#   cat("4. Group by 'city' to compare Cibolo vs Schertz\n")
#   cat("5. Use 'category' to group related indicators\n\n")
# } else {
#   cat("=== SHINY STANDARDIZATION SKIPPED ===\n")
#   cat("Set STANDARDIZE_FOR_SHINY = TRUE to create Shiny-ready datasets\n\n")
# }

```


#==============================================================================
# STEP 8: CREATE CHARTS AND ANALYSIS
#==============================================================================

```{r}

# Key indicators for automatic chart generation
key_indicators <- c(
  "Total Population",
  "Median Household Income", 
  "Per Capita Income",
  "Median Home Value",
  "Percent Bachelor's Degree+",
  "Unemployment Rate (%)"
)

# Create dashboard of key charts
key_charts <- create_chart_dashboard(longitudinal_data, key_indicators, save_plots = TRUE)

# Display charts in R
cat("=== DISPLAYING KEY CHARTS ===\n")
walk2(key_charts, names(key_charts), function(chart, name) {
  if(!is.null(chart)) {
    print(chart)
    cat("\n--- Chart for", name, "displayed ---\n\n")
  }
})

# Enhanced Per Capita Income Analysis (as requested)
cat("=== DETAILED PER CAPITA INCOME ANALYSIS ===\n")
per_capita_analysis <- analyze_variable_trends(longitudinal_data, "Per Capita Income")
per_capita_chart <- create_line_chart(longitudinal_data, "Per Capita Income")

if(!is.null(per_capita_chart)) {
  print(per_capita_chart)
  
  year_range <- paste0(min(ANALYSIS_YEARS), "_", max(ANALYSIS_YEARS))
  ggsave(paste0("per_capita_income_detailed_", year_range, ".png"), 
         per_capita_chart, width = 12, height = 8, dpi = 300)
  cat("Detailed Per Capita Income chart saved.\n\n")
}

```


#==============================================================================
# STEP 9: EXPORT ANALYSIS DATA
#==============================================================================

```{r}

# Create filename with year range
year_range <- paste0(min(ANALYSIS_YEARS), "_", max(ANALYSIS_YEARS))

# Export data in multiple formats
chart_ready_data <- longitudinal_data %>%
  select(city, year, period, category, variable_label, estimate, moe) %>%
  arrange(category, variable_label, city, year)

# # Excel workbook with multiple sheets
# write_xlsx(
#   list(
#     "Chart_Ready_Data" = chart_ready_data,
#     "Wide_Format" = longitudinal_wide,
#     "Cibolo_Only" = cibolo_data,
#     "Schertz_Only" = schertz_data,
#     "Variable_List" = tibble(
#       variable_code = names(longitudinal_vars),
#       variable_label = unname(longitudinal_vars),
#       description = "ACS 5-Year Estimates"
#     )
#   ),
#   path = paste0("cibolo_schertz_analysis_", year_range, ".xlsx")
# )
# 
# # CSV exports for analysis data
# write_csv(longitudinal_data, paste0("longitudinal_analysis_data_", year_range, ".csv"))
# write_csv(longitudinal_wide, paste0("longitudinal_wide_format_", year_range, ".csv"))

```


#==============================================================================
# STEP 10: SUMMARY AND USER FUNCTIONS
#==============================================================================

```{r}

cat("=== ENHANCED CENSUS ANALYSIS SUMMARY ===\n")
cat("Analysis period:", min(ANALYSIS_YEARS), "to", max(ANALYSIS_YEARS), "\n")
cat("Total years analyzed:", length(unique(longitudinal_data$year)), "\n")
cat("Variables tracked:", length(longitudinal_vars), "\n")
cat("Charts created:", length(key_charts), "\n")
cat("Analysis files saved: cibolo_schertz_analysis_", year_range, ".xlsx\n")

if(STANDARDIZE_FOR_SHINY) {
  cat("Shiny standardization: ENABLED (files will be saved at end)\n")
} else {
  cat("Shiny standardization: DISABLED\n") 
}

cat("\n")

# Show available variables for plotting
cat("=== AVAILABLE VARIABLES FOR LINE CHARTS ===\n")
available_vars <- longitudinal_data %>%
  distinct(category, variable_label) %>%
  arrange(category, variable_label)

for(cat_name in unique(available_vars$category)) {
  cat("\n", cat_name, ":\n")
  vars_in_cat <- available_vars %>% filter(category == cat_name) %>% pull(variable_label)
  cat(paste("-", vars_in_cat), sep = "\n")
}

# Enhanced user function for creating charts
plot_comparison <- function(variable_name, save_plot = FALSE, width = 12, height = 8, analyze_trends = FALSE) {
  chart <- create_line_chart(longitudinal_data, variable_name)
  
  if(!is.null(chart)) {
    print(chart)
    
    if(save_plot) {
      year_range <- paste0(min(ANALYSIS_YEARS), "_", max(ANALYSIS_YEARS))
      filename <- paste0(str_replace_all(str_to_lower(variable_name), "[^a-z0-9]", "_"), 
                        "_", year_range, ".png")
      ggsave(filename, chart, width = width, height = height, dpi = 300)
      cat("Chart saved as:", filename, "\n")
    }
    
    if(analyze_trends) {
      analyze_variable_trends(longitudinal_data, variable_name)
    }
  }
  
  return(chart)
}

# Function to quickly change years and re-run analysis
rerun_analysis <- function(new_years, key_vars = key_indicators, shiny_standardize = STANDARDIZE_FOR_SHINY) {
  cat("=== RE-RUNNING ANALYSIS WITH NEW YEARS ===\n")
  cat("New years:", paste(new_years, collapse = ", "), "\n\n")
  
  # Update global variable
  ANALYSIS_YEARS <<- new_years
  
  # Re-run data collection
  new_data <- get_longitudinal_data(years = new_years)
  
  # Create new charts
  new_charts <- create_chart_dashboard(new_data, key_vars, save_plots = TRUE)
  
  # Standardize for Shiny if requested
  if(shiny_standardize) {
    new_shiny_data <- standardize_for_shiny_app(new_data)
    save_shiny_data(new_shiny_data, "cibolo_schertz_census_rerun")
  }
  
  # Quick export
  year_range <- paste0(min(new_years), "_", max(new_years))
  write_csv(new_data, paste0("rerun_analysis_data_", year_range, ".csv"))
  
  cat("Analysis rerun complete!\n")
  return(new_data)
}

# Growth summary for all key variables
growth_summary <- longitudinal_data %>%
  filter(variable_label %in% key_indicators) %>%
  group_by(city, variable_label) %>%
  summarise(
    start_value = first(estimate, order_by = year),
    end_value = last(estimate, order_by = year),
    total_growth_pct = round(((end_value / start_value) - 1) * 100, 1),
    annual_growth_rate = round((((end_value / start_value)^(1/(max(year) - min(year)))) - 1) * 100, 2),
    .groups = "drop"
  ) %>%
  arrange(variable_label, city)

cat("\n=== GROWTH SUMMARY FOR KEY INDICATORS ===\n")
print(growth_summary)

cat("\n=== USAGE EXAMPLES ===\n")
cat("Create any line chart:\n")
cat("plot_comparison('Total Population')\n")
cat("plot_comparison('Median Home Value', save_plot = TRUE, analyze_trends = TRUE)\n\n")

cat("Change analysis years and re-run:\n")
cat("rerun_analysis(2015:2022)  # Recent years only\n")
cat("rerun_analysis(c(2009, 2015, 2022))  # Snapshot comparison\n")
cat("rerun_analysis(2010:2019)  # Pre-pandemic decade\n\n")

cat("Toggle Shiny standardization:\n")
cat("STANDARDIZE_FOR_SHINY <- FALSE  # Disable Shiny output\n")
cat("rerun_analysis(2015:2022, shiny_standardize = TRUE)  # Force Shiny output\n")
```


#==============================================================================
# FINAL STEP: EXPORT ALL FILES (CONDITIONAL)
#==============================================================================

```{r}

# Create year range for filenames
year_range <- paste0(min(ANALYSIS_YEARS), "_", max(ANALYSIS_YEARS))

# 1. Standard analysis files (always created)
write_xlsx(
  list(
    "Chart_Ready_Data" = chart_ready_data,
    "Wide_Format" = longitudinal_wide,
    "Cibolo_Only" = cibolo_data,
    "Schertz_Only" = schertz_data,
    "Variable_List" = tibble(
      variable_code = names(longitudinal_vars),
      variable_label = unname(longitudinal_vars),
      description = "ACS 5-Year Estimates"
    )
  ),
  path = paste0("cibolo_schertz_analysis_", year_range, ".xlsx")
)

# 2. Shiny-standardized files (conditional)
if(STANDARDIZE_FOR_SHINY) {
  cat("Creating Shiny-standardized files...\n")
  
  # Create the standardized data: date, value, + categories
  shiny_standardized_data <- standardize_for_shiny_app(longitudinal_data)
  
  # Save in both formats
  shiny_files <- save_shiny_data(shiny_standardized_data, "cibolo_schertz_census")
  
  cat("Shiny files created:", shiny_files$csv, "and", shiny_files$rds, "\n")
}

cat("=== ALL FILE EXPORTS COMPLETE ===\n")

```



## Analysis Summary

This enhanced analysis provides:

### **Data Collection:**
- **Maximum 14 years** of longitudinal data (2009-2022)
- **Comprehensive variables** across demographics, economics, education, and housing
- **Data availability checking** before processing

### **Visualization:**
- **Professional line charts** with automatic formatting
- **Chart dashboard** creation and export
- **Trend analysis** with growth rates and comparative statistics

### **Shiny Integration:**
- **Optional standardization** with `STANDARDIZE_FOR_SHINY` toggle
- **Standardized format:** `date`, `value`, + categorical columns
- **Dual file output:** Both .csv (universal) and .rds (R-optimized)
- **Single combined file** with all variables and cities

### **Flexibility:**
- **User-defined years** at the top of the script
- **Quick rerun functions** for different time periods
- **Multiple export formats** for different use cases

The standardized datasets are ready for immediate upload to your Shiny dashboard, with `date` and `value` as the first two columns as required, while preserving all categorical information for filtering and grouping in your app's line chart functionality.

