---
title: "PUMS Educ Crosstabs"
subtitle: ""
description: ""
authors: 
  - name: "Dan Swart, CPA (ret)"
  - name: "Claude Sonnet 4.5"
date: today
date-format: long
# bibliography: manual-refs.bib
format:
  html:
    resources:
      - reference-backlinks.js
    include-after-body:    
      - text: |
          # <script type="text/javascript" src="reference-backlinks.js"></script>
    default: true         
    code-copy: true
    code-link: true        # This adds individual buttons
    code-fold: true        # Hide code by default, show on click
    code-summary: "Show the code"
    code-overflow: wrap
    code-block-bg: "#FAEBD7"
    code-block-border-left: "#31BAE9"
    embed-resources: false     # Fast for drafts. Override for sharing output
    include-in-header:
      - text: 
          <link href="https://fonts.googleapis.com/css2?family=Lato&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Fira+Mono&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Montserrat&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Cabin&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Roboto+Slab&display=swap" rel="stylesheet">
          <link href="https://fonts.googleapis.com/css2?family=Schoolbell&display=swap" rel="stylesheet">
      - header.html
    css:
      - swart.css
      - tachyons.min.css
      - r-colors.css
    fontsize: 18pt
    lightbox: true
    page-layout: full
    fig-width: 12
    fig-height: 10
    fig-dpi: 300
    html-math-method: katex
    df-print: paged
    toc: true
    toc-float: true
    citeproc: true
    link-citations: true
    linestretch: 1.0
    
    
    
  typst:
    fig-width: 12
    fig-height: 10
    fig-dpi: 300
    margin:
      x: 1in
      y: 1in
    toc: true
    fontsize: 16pt
    mainfont: "Cabin"
  

    
  revealjs:
    slide-number: true
    transition: fade
    code-overflow: wrap
    center: true
    smaller: true
    scrollable: true
    chalkboard: true
    multiplex: false
    theme: solarized
    reference-location: margin
    logo: img/red-cross-640-435.png
    footer: "Footer text"
    code-block-height: 650px



  # docx:
  #   highlight-style: github
  #   fig_caption: true



editor: source

quarto:
  render:
    cache-refresh: true


# for .qmd filesd
execute:
  echo: true
  message: false
  warning: false
  eval: true
  fig-width: 12
  fig-height: 10


# for .rmd files
knitr:
  opts_chunk:
    echo: true
    error: false
    warning: false
    message: false
    cache: false


---


```{r}
#| label: setup
#| include: false


# install.packages(c("mapview", "survey", "srvyr", "arcgislayers"))

# census_api_key("95496766c51541ee6f402c1e1a8658581285b759", install = TRUE, overwrite = TRUE)


# # load libraries - NOT NEEDED


# Force dplyr's select to take precedence
select <- dplyr::select
filter <- dplyr::filter

# Options
options(scipen = 999)
options(qic.clshade = T) # NO LONGER NEEDED; CHARTS ALL PREPARED WITH GGPLOT2 ONLY
options(qic.linecol = 'black') # NO LONGER NEEDED; CHARTS ALL PREPARED WITH GGPLOT2 ONLY
options(qic.signalcol = "firebrick") # NO LONGER NEEDED; CHARTS ALL PREPARED WITH GGPLOT2 ONLY
options(qic.targetcol = "purple") # NO LONGER NEEDED; CHARTS ALL PREPARED WITH GGPLOT2 ONLY
options(DT.options = list(dom = 'pBlfrti')) # Add buttons, filtering, and top (t) pagination controls
options(shiny.maxRequestSize = 50 * 1024^2) # Set upload maximum to 50 MB
options(tigris_use_cache = TRUE)
options(device = "RStudioGD") 


# Flextable defaults:
flextable::set_flextable_defaults(
  font.size = 14, 
  font.family = "Cabin",
  font.color = "black",
  table.layout = "fixed",
  border.color = "darkgray",
  padding.top = 3, padding.bottom = 3,
  padding.left = 4, padding.right = 4,
  line_spacing = 1.3,
  digits = 2,
  decimal.mark = ".",
  big.mark = " ",
  na_str = "<na>",
  post_process_html = identity,
  post_process_docx = identity
)


#
# # Sample Code:
# flextable::flextable(violations) |>
 #  flextable::set_header_labels(
 #    Variable = "Variable",
 #    Measurement = "Measurement",
 #    Likely_Impact = "Likely Impact"
 #  ) |>
#    flextable::add_header_lines(values = "Frequent Violations of Scientific Method in Current So-Called 'Equity' Research") |>
#   flextable::color(i = 1, color = "blue", part = "header") |>
#   flextable::italic(i = 1, part = "header") |>
#   flextable::align(i = 1, align = "center", part = "header") |>
#   flextable::fontsize(i = 1, size = 14, part = "header") |>
#   flextable::bg(i = 1, bg = "white", part = "header") |>
#   flextable::bg(i = 2, bg = "palegreen", part = "header") |>
#   flextable::bold(i = 1:2, part = "header") |>
#   flextable::bold(i = 1:7, j = 1, part = "body") |>
#   ftExtra::colformat_md() |> 
#   flextable::autofit()
  

# Flextable built-in themes:
  # flextable::theme_alafoli()	|>  # BLAH
  # flextable::theme_apa()  # THIS IS NICE
  # flextable::theme_booktabs() |>  # NICE, MORE COMPACT
  # flextable::theme_box() |>   # OK, INCLUDES CELL BORDERS
  # flextable::theme_tron() |>  # 'DARK MODE' BLUE TEXT
  # flextable::theme_tron_legacy() |>   # 'DARK MODE' YELLOW TEXT
  # flextable::theme_vader() |>    # 'DARK MODE' WHITE TEXT
  # flextable::theme_vanilla() |>   # NOT SPECIAL
  # flextable::theme_zebra()	|>
  #

# Flextable titles:
  # flextable::colformat_double(j = c("Mean", "SD", "N"), big.mark = ",", digits = 1) |>
  # flextable::flextable(variance_comparison) |>
  #  flextable::add_header_lines(values = "The Within-Group vs Between-Group Variance Problem") |>

# Flextable Title theming at table-level:
  #  flextable::color(i = 1, color = "blue", part = "header") |>
  #  flextable::italic(i = 1, part = "header") |>
  #  flextable::align(i = 1, align = "center", part = "header") |>
  #  flextable::fontsize(i = 1, size = 14, part = "header") |>
  #  flextable::bg(i = 1, bg = "white", part = "header") |>

# Flextable standard table background colors:
  #  flextable::bg(i = 2, bg = "palegreen", part = "header") |>
  

# Flextable reading markdown:
  #  ftExtra::colformat_md() |> 

# Flextable auto-sizing cell widths:
  #  flextable::autofit() 
 
 # Flextable background based on SPECIFIC cell contents:
  #
  # flextable::bg(i = ~ Impact_on_Validity == "High", j = "Impact_on_Validity", bg = "#ffcccc") |>
 # flextable::bg(i = ~ Impact_on_Validity == "Medium", j = "Impact_on_Validity", bg = "#ffffcc") |>
  #
  # 
  # Apply yellow background to any cell containing "Yes":

  # for (col in base::names(hypotheses_data)) {
  #   yes_rows <- base::which(hypotheses_data[[col]] == "Yes")
  #   if (base::length(yes_rows) > 0) {
  #     ft <- ft |>
  #       flextable::bg(i = yes_rows, j = col, bg = "yellow", part = "body")
  #   }
  # } 
  #
  # Apply to last row of table:
  #
  #



# Set global theme for consistent plots
ggplot2::theme_set(
  ggplot2::theme_minimal(base_size = 20) +
    ggplot2::theme(
      plot.title.position = "plot",
      plot.title = ggtext::element_textbox_simple(
        family = "Cabin",
        face = "bold",
        color = "darkgreen",
        size = 26,
        fill = "yellow",
        lineheight = 1.2,
        padding = ggplot2::margin(5.5, 5.5, 0.0, 5.5),
        margin = ggplot2::margin(0, 0, 5.5, 0)
      ),
      plot.subtitle = ggtext::element_textbox_simple(
        family = "Cabin",
        color = "darkgreen",
        face = "bold",
        size = 24,
        fill = "yellow",
        lineheight = 1.2,
        padding = ggplot2::margin(5.5, 5.5, 5.5, 5.5),
        margin = ggplot2::margin(10.5, 0, 5.5, 0)
      ),
      plot.caption = ggtext::element_markdown(
        family = "Cabin",
        size = 22,
        hjust = 1,
        color = "darkblue",
        face = "italic",
        fill = "yellow",
        lineheight = 1.0
      ),
      axis.text.x = ggtext::element_markdown(
        family = "Cabin",
        face = "bold",
        color = "blue",
        size = 20,
        angle = 45,
        hjust = 1
      ),
        # ggplot2::element_blank(),
      axis.title.x = ggtext::element_markdown(
        family = "Cabin",
        face = "bold",
        color = "blue",
        size = 20),
        # ggplot2::element_blank(),
      axis.text.y = ggtext::element_markdown(
        family = "Cabin",
        face = "bold",
        color = "blue",
        size = 20,
        angle = 45,
        hjust = 1
      ),
        # ggplot2::element_blank(),
      axis.title.y = ggtext::element_markdown(
        family = "Cabin",
        face = "bold",
        color = "blue",
        size = 20),
        # ggplot2::element_blank(),
      strip.text = ggtext::element_markdown(
        family = "Cabin",
        color = "black",
        size = ggplot2::rel(1.1),
        face = "italic",
        margin = ggplot2::margin(2, 0, 0.5, 0, "lines")
      ),
      axis.text = ggtext::element_markdown(
        family = "Cabin",
        color = "black"),
      panel.background = ggplot2::element_rect(fill = "white", color = NA),
      plot.background = ggplot2::element_rect(fill = "white", color = NA),
      legend.position = "none",
      panel.spacing.x = grid::unit(1.5, "cm"),
      panel.spacing.y = grid::unit(1.5, "cm"),
      plot.margin = ggplot2::margin(20, 20, 20, 20, "pt")
    )
)



  
# Set seed for reproducibility
base::set.seed(123)

```


# PUMAs for SCUC-ISD      
For your PUMS analysis, use these PUMA codes:
  
  # Define the PUMAs that cover SCUC ISD area
  scuc_pumas <- c("05700", "05916")  # Remove the "48" state prefix for PUMS data
  
  # Filter your education data for these PUMAs
  scuc_education_data <- education_data %>%
    dplyr::filter(PUMA %in% scuc_pumas)
  
  # Or if you want to include surrounding areas for comparison:
  extended_area_pumas <- c(
    "05700",  # Guadalupe County (Schertz, Cibolo)
    "05916",  # Bexar County Northeast (Universal City)
    "05800",  # Comal County (adjacent northern area)
    "05915"   # Bexar County North (adjacent area)
  )
  
  cat("SCUC ISD area PUMAs: 05700 (Guadalupe County), 05916 (Bexar County Northeast)\n")




```{r}
#| label: load libraries

library(tidyverse)
library(tidycensus)

```


STEP 1: Define your comparison groups and outcomes


# STEP 1: Use available years and variables

```{r}

working_years <- 2017:2022  # 6 years of data
working_vars <- c("SCHL", "FOD1P", "FOD2P", "RAC1P", "HISP", "AGEP", "SEX",
                  "HINCP", "POVPIP", "NP", "HHT", "MAR", "WAGP", "WKHP", "ESR")
# Note: Removed NPOC since it doesn't exist

cat("Downloading PUMS data for", length(working_years), "years with", length(working_vars), "variables...\n")

```

# STEP 2: Download and standardize data types for each year (this will take a while)

```{r}

# Download and standardize only existing columns
education_data_list <- purrr::map(working_years, ~{
  cat("Downloading year", .x, "...\n")
  
  year_data <- tidycensus::get_pums(
    variables = working_vars,
    state = "TX",
    year = .x,
    survey = "acs5"
  ) 
  
  # Check what columns we actually got
  cat("  Columns received:", paste(names(year_data), collapse = ", "), "\n")
  
  # Add year and standardize only existing columns
  year_data <- year_data %>%
    dplyr::mutate(
      year = .x,
      
      # Standardize columns that exist in our requested variables
      AGEP = as.numeric(AGEP),
      HINCP = as.numeric(HINCP),
      WAGP = as.numeric(WAGP),
      WKHP = as.numeric(WKHP),
      POVPIP = as.numeric(POVPIP),
      NP = as.numeric(NP),
      
      # Character variables from our list
      RAC1P = as.character(RAC1P),
      HISP = as.character(HISP),
      SEX = as.character(SEX),
      MAR = as.character(MAR),
      ESR = as.character(ESR),
      HHT = as.character(HHT),
      SCHL = as.character(SCHL),
      FOD1P = as.character(FOD1P),
      FOD2P = as.character(FOD2P)
    )
  
  # Standardize automatic columns if they exist
  if("PWGTP" %in% names(year_data)) {
    year_data$PWGTP <- as.numeric(year_data$PWGTP)
  }
  if("SPORDER" %in% names(year_data)) {
    year_data$SPORDER <- as.character(year_data$SPORDER)
  }
  if("PUMA" %in% names(year_data)) {
    year_data$PUMA <- as.character(year_data$PUMA)
  }
  if("SERIALNO" %in% names(year_data)) {
    year_data$SERIALNO <- as.character(year_data$SERIALNO)
  }
  if("ST" %in% names(year_data)) {
    year_data$ST <- as.character(year_data$ST)
  }
  
  return(year_data)
})

# Combine the datasets
education_data <- dplyr::bind_rows(education_data_list)

cat("Download complete! Data dimensions:", dim(education_data), "\n")

```


STEP 3: Create analysis-ready dataset 

```{r}

# Create analysis-ready dataset with proxy for children
analysis_ready <- education_data %>%
  dplyr::mutate(
    # Define racial/ethnic groups using CORRECT character comparisons
    race_ethnicity = dplyr::case_when(
      RAC1P == "1" & HISP == "01" ~ "White_NH",      # White, not Hispanic
      RAC1P == "2" & HISP == "01" ~ "Black_NH",      # Black, not Hispanic
      RAC1P == "6" & HISP == "01" ~ "Asian_NH",      # Asian, not Hispanic
      HISP == "02" ~ "Hispanic",                     # Hispanic (any race)
      TRUE ~ "Other"
    ),
    
    # Convert SCHL to numeric (handles leading zeros automatically)
    SCHL_num = as.numeric(SCHL),  # "21" becomes 21, "bb" becomes NA
    college_plus = ifelse(SCHL_num >= 21, 1, 0),
    graduate_degree = ifelse(SCHL_num >= 22, 1, 0),
    
    # Demographics and controls
    age_group = cut(AGEP, breaks = c(0, 25, 35, 45, 55, 65, 100)),
    married = ifelse(MAR == "01", 1, 0),             # Married is probably "01"
    income_quintile = dplyr::ntile(HINCP, 5),
    full_time = ifelse(WKHP >= 35, 1, 0),
    
    # Proxy for having children - need to check HHT codes
    likely_has_children = dplyr::case_when(
      HHT %in% c("01", "02") & NP >= 3 ~ 1,         # Family households with 3+ people
      HHT %in% c("03", "04") & NP >= 2 ~ 1,  
      TRUE ~ 0
    )
  ) %>%
  dplyr::filter(
    AGEP >= 25 & AGEP <= 65,  # Working age adults
    race_ethnicity != "Other",
    !is.na(SCHL_num)  # Remove records with invalid education codes
  )

cat("Fixed analysis dataset dimensions:", dim(analysis_ready), "\n")

# Check the race/ethnicity distribution
cat("Race/ethnicity distribution:\n")
print(table(analysis_ready$race_ethnicity, useNA = "always"))

```



STEP 4: Example hypothesis test

```{r}

# Hypothesis 1: Racial achievement gaps controlling for socioeconomic factors
hypothesis_1 <- analysis_ready %>%
  dplyr::filter(AGEP >= 30 & AGEP <= 40) %>%  # Peak career-building years
  dplyr::group_by(year, race_ethnicity, income_quintile, likely_has_children, married) %>%
  dplyr::summarise(
    college_plus_rate = stats::weighted.mean(college_plus, PWGTP, na.rm = TRUE),
    graduate_degree_rate = stats::weighted.mean(graduate_degree, PWGTP, na.rm = TRUE),
    n = sum(PWGTP), 
    .groups = "drop"
  ) %>%
  dplyr::filter(n >= 30)  # Minimum sample size for reliable estimates

# Create the line graph test
ggplot2::ggplot(hypothesis_1, ggplot2::aes(x = year, y = college_plus_rate, color = race_ethnicity)) +
  ggplot2::geom_line(size = 1.2) +
  ggplot2::facet_wrap(~paste("Income Q", income_quintile, "| Children:", likely_has_children)) +
  ggplot2::labs(
    title = "College Completion Rates: Apples-to-Apples Comparison",
    subtitle = "Controlling for income and family structure",
    y = "College Completion Rate",
    caption = "If gaps persist within same income/family groups, socioeconomic factors don't explain them"
  ) +
  ggplot2::theme_minimal()

```





